Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the `eechidna` `R` package.

### 2019-03-31 draft-paper.Rmd
add modelling intro bit

*1: HERE Introductory paragraph*

## Data pre-processing
Having used the imputation method detailed in Section \@ref(data), all socio-demographic variables are not for each electorate in each election, and are joined to with the corresponding two-party preferred vote data. Socio-demographic variables within each election year are standardized to have mean zero and variance one, to adjust for changing variable scales. For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016. By standardizing, the median rental price of an electorate is reported relative to all other electorates in the same year.

@@

Socio-demographic variables[^2] that represent similar information are combined into "factors", informed by principal component analysis (PCA) on a combined dataset of socio-demographic variables across the six elections. A factor is created by combining several variables that have large loadings in a particular component and an intuitive justification for their grouping. Only the first four principal components are considered (noting that the scree plot levels off after the fourth component). A loading with magnitude greater than 0.15 is considered large. After computing these *sums*, each factor is again standardized to have mean zero and variance one, within each election. 
*2: HERE Explain the different factors and how they are constructed - also justify combining all cross-sections*

>>>>

*1: HERE Introductory paragraph*
Electoral socio-demographic variables are now available for each of the six elections and are joined with their corresponding two-party preferred votes. Before choosing an appropriate model, two issues with the socio-demographic variables are addressed. Firstly, variable scales change over the years and as such the data is standardized accordingly. Secondly, many variables represent similar information and where appropriate, should be combined in some way. Principal component analysis is used to determine which intuitive groups of variables covary and should be combined into a single variable. This also reduces the dimension of the data. After these steps, a model specification is chosen.

## Data pre-processing
Many of the socio-demographic variables have changing scales over the years. For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016. In order for socio-demographic effects to be comparable across years, the explanatory variables are standardized to have mean zero and variance one within each election year. By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.

@@

Socio-demographic variables[^2] that represent similar information are combined into "factors", informed by principal component analysis (PCA) on a combined dataset of socio-demographic variables across the six elections. A factor is created by combining several variables that have large loadings in a particular component and an intuitive justification for their grouping. Only the first four principal components are considered (noting that the scree plot levels off after the fourth component). A loading with magnitude greater than 0.15 is considered large. After computing these *sums*, each factor is again standardized to have mean zero and variance one, within each election.

*2: HERE Explain the different factors and how they are constructed - also justify combining all cross-sections*
